Lesson 4
========================================================
library(ggplot2)
pf = read.csv("", sep = \t)
***

### Scatterplots and Perceived Audience Size
Notes:

***

### Scatterplots
Notes:

```{r Scatterplots}

ggplot(data = pf, aes(x = age, y = friend_count)) +
  geom_point(alpha = 1/10, position = position_jitter(h=0)) +
  xlim(13, 90) +
  coord_trans(y = "sqrt")
```

***

#### What are some things that you notice right away?
Response:

***

### ggplot Syntax
Notes:

```{r ggplot Syntax}

```

***

### Overplotting
Notes:

```{r Overplotting}

```

#### What do you notice in the plot?
Response:

***

### Coord_trans()
Notes:

```{r Coord_trans()}

```

#### Look up the documentation for coord_trans() and add a layer to the plot that transforms friend_count using the square root function. Create your plot!

```{r}

```

#### What do you notice?

***

### Alpha and Jitter
Notes: friend_initiated, age

```{r Alpha and Jitter}
ggplot(data = pf, aes(x = age, y = friendships_initiated)) + 
  geom_point(alpha = 1/15, position = position_jitter(h=0))


```

***

### Overplotting and Domain Knowledge
Notes:

***

### Conditional Means
Notes:

```{r Conditional Means}
pf.age <- pf %>% 
  group_by(age) %>% 
  summarise(friend_count_mean = mean(friend_count), 
            friend_count_median = median(friend_count),
            n = n()) 


```

Create your plot!

```{r Conditional Means Plot}
ggplot(data = pf.age, aes(x = age, y = friend_count_mean)) + 
  geom_point()


```

***

### Overlaying Summaries with Raw Data
Notes:

```{r Overlaying Summaries with Raw Data}
library(ggplot2)
ggplot(data = pf, aes(x = age, y = friend_count)) + 
  geom_point(alpha = 1/20, position = position_jitter(h=0), color = 'orange') + 
  xlim(13, 90) + 
  geom_line(stat = 'summary', fun.y = mean ) + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9), linetype = 2, color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.1), linetype =2,
            color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.5), color = 'blue')
  
ggplot(data = pf, aes(x = age, y = friend_count)) + 
  geom_point(alpha = 1/20, position = position_jitter(h=0), color = 'orange') + 
  coord_cartesian(xlim = c(13, 90), ylim = c(0, 1000)) +
  geom_line(stat = 'summary', fun.y = mean ) + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.9), linetype = 2, color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.1), linetype =2,
            color = 'blue') + 
  geom_line(stat = 'summary', fun.y = quantile, fun.args = list(probs = 0.5), color = 'blue')
    


```

#### What are some of your observations of the plot?
Response:

***

### Moira: Histogram Summary and Scatterplot
See the Instructor Notes of this video to download Moira's paper on perceived audience size and to see the final plot.

Notes:

***

### Correlation
Notes:

```{r Correlation}



```

Look up the documentation for the cor.test function.

What's the correlation between age and friend count? Round to three decimal places.
Response:

***

### Correlation on Subsets
Notes:

```{r Correlation on Subsets}
with(                 , cor.test(age, friend_count))
```

***

### Correlation Methods
Notes:

***

## Create Scatterplots
Notes:

```{r}

```

***

### Strong Correlations
Notes:
# Create a scatterplot of likes_received (y)
# vs. www_likes_received (x)

```{r Strong Correlations}
ggplot(data = pf, aes(x = www_likes_received, y = likes_received)) +
  geom_point() +
  xlim(0, quantile(pf$www_likes_received, 0.95)) + 
  ylim(0, quantile(pf$likes_received, 0.95)) + 
  geom_smooth(method = 'lm', color = 'blue')

with(pf, cor.test(www_likes_received, likes_received))

```

What's the correlation betwen the two variables? Include the top 5% of values for the variable in the calculation and round to 3 decimal places.

```{r Correlation Calcuation}

```

Response:

***

### Moira on Correlation
Notes:

***

### More Caution with Correlation
Notes:

```{r More Caution With Correlation}
install.packages('alr3')
library(alr3)
```

Create your plot!

```{r Temp vs Month}
data("Mitchell")
?Mitchell
ggplot(data = Mitchell, aes(x = Month, y = Temp)) + 
  geom_point() + 
  scale_x_continuous(breaks = seq(0, 203, 12))

```

***

### Noisy Scatterplots
a. Take a guess for the correlation coefficient for the scatterplot.

b. What is the actual correlation of the two variables?
(Round to the thousandths place)

```{r Noisy Scatterplots}

```

***

### Making Sense of Data
Notes:

```{r Making Sense of Data}

```

***

### A New Perspective

What do you notice?
Response:

Watch the solution video and check out the Instructor Notes!
Notes:

***

### Understanding Noise: Age to Age Months
Notes:

```{r Understanding Noise: Age to Age Months}

pf$age_with_months <- pf$age + (1 - pf$dob_month / 12)

pf.fc_by_age_months <- pf %>% 
  group_by(age_with_months) %>% 
  summarise(friend_count_mean = mean(friend_count), 
            friend_count_median = mean(friend_count),
            n =n())

ggplot(data = pf.fc_by_age_months, aes(x = age_with_months, y = friend_count_mean)) + 
  geom_line() + 
  xlim(0, 71)

```

***

### Age with Months Means

```{r Age with Months Means}



```

Programming Assignment
```{r Programming Assignment}

```

***

### Noise in Conditional Means

```{r Noise in Conditional Means}

```

***

### Smoothing Conditional Means
Notes:

```{r Smoothing Conditional Means}

```

***

### Which Plot to Choose?
Notes:

***

### Analyzing Two Variables
Reflection:

we learn how to explore the relationship between two variables. Our main visulization tool is scatterplot. But we also augmented the scatterplot, with conditional summaries, like means, 
we also learn the benefits and limitation of using correlation. To understand the relationship between two variables and how correlation may effect your decisions over which variables to include in your final models. Another, we learn hwo make sense of data through adjusting our visulizations. We learned not to necesarrily trust our interpretation of initial scatterplots like with the seasonal temperature data. And we learned how to use jitter and transparency to reduce over plotting 

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes! 

