Lesson 5
========================================================

### Multivariate Data
Notes:

***

### Moira Perceived Audience Size Colored by Age
Notes:

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_histogram()

# Write code to create a new data frame,
# called 'pf.fc_by_age_gender', that contains
# information on each age AND gender group.

# The data frame should contain the following variables:

#    mean_friend_count,
#    median_friend_count,
#    n (the number of users in each age and gender grouping)

# Here is an example of the structure of your data frame. Your
# data values will be different. Note that if you are grouping by
# more than one variable, you will probably need to call the
# ungroup() function. 

#   age gender mean_friend_count median_friend_count    n
# 1  13 female          247.2953                 150  207
# 2  13   male          184.2342                  61  265
# 3  14 female          329.1938                 245  834
# 4  14   male          157.1204                  88 1201

pf.fc_by_age_gender <- pf %>%
  group_by(age, gender) %>% 
  summarise(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %>% 
  arrange(age)
  



```



***

### Plotting Conditional Summaries
Notes:
# Create a line graph showing the
# median friend count over the ages
# for each gender. Be sure to use
# the data frame you just created,
# pf.fc_by_age_gender.
```{r Plotting Conditional Summaries}
ggplot(subset(pf.fc_by_age_gender, !is.na(gender)), aes(x = age, y = mean_friend_count)) + 
  geom_line(aes(color = gender))


```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes:

***

### Reshaping Data
Notes:

```{r}
install.packages('reshape2')
library(reshape2)
spread_pf <- subset(pf.fc_by_age_gender, !is.na(gender))[c('age', 'gender', 'median_friend_count')] %>% 
  spread(gender, median_friend_count) %>% 
  mutate(ratio = male / female)

ggplot(data = spread_pf, aes(x = age, y = ratio)) + 
  geom_line() + 
  geom_hline(aes(yintercept=1), linetype = 'dashed', color = 'blue')


```


***

### Ratio Plot
Notes:

```{r Ratio Plot}

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
# Create a variable called year_joined
# in the pf data frame using the variable
# tenure and 2014 as the reference year.

# The variable year joined should contain the year
# that a user joined facebook.

pf$year_joined <- floor(2014 - pf$tenure / 365)

```

***

### Cut a Variable
Notes:
# Create a new variable in the data frame
# called year_joined.bucket by using
# the cut function on the variable year_joined.

# You need to create the following buckets for the
# new variable, year_joined.bucket

#        (2004, 2009]
#        (2009, 2011]
#        (2011, 2012]
#        (2012, 2014]

# Note that a parenthesis means exclude the year and a
# bracket means include the year.

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))

ggplot(data = subset(pf, !is.na(gender)), aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = median)
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
pf$year_joined.bucket <- cut(pf$year_joined, c(2004, 2009, 2011, 2012, 2014))

ggplot(data = subset(pf, !is.na(year_joined.bucket)), aes(x = age, y = friend_count)) +
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) + 
  geom_line(stat = 'summary', fun.y = mean, linetype = 2)
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}

```

***

### Friending Rate
Notes:

```{r Friending Rate}
sub <- subset(pf, pf$tenure >= 1)
summary(sub$friend_count/sub$tenure)




```

***

### Friendships Initiated
Notes:

What is the median friend rate?

What is the maximum friend rate?

```{r Friendships Initiated}
# Create a line graph of mean of friendships_initiated per day (of tenure)
# vs. tenure colored by year_joined.bucket.

# You need to make use of the variables tenure,
# friendships_initiated, and year_joined.bucket.

# You also need to subset the data to only consider user with at least
# one day of tenure.
ggplot(data = sub, aes(x = tenure, y = friendships_initiated / tenure)) + 
  geom_line(aes(color = year_joined.bucket), stat = 'summary', fun.y = mean) 
  

```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}
library(gridExtra)
p1 <- ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

p2 <- ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p3 <- ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

p4 <- ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

grid.arrange(p1, p2, p3, p4, ncol = 1)
```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv('yogurt.csv')
yo$id <- factor(yo$id)

ggplot(data = yo, aes(x = price)) + 
  geom_histogram() +
  scale_x_continuous(limits = c(20, 70), breaks = seq(20,70,5))
```

***

### Number of Purchases
Notes: tansform(dataframe, column)

```{r Number of Purchases}
yo <- transform(yo, all.purchase = strawberry + blueberry + pina.colada + plain + mixed.berry)
qplot(data = yo, x = all.purchase, binwidth = 1)
```


***

### Prices over Time
Notes:

```{r Prices over Time}

ggplot(data = yo, aes(x = time, y = price)) + 
  geom_point(alpha = 1/20, shape = 21)
  

```

***

### Sampling Observations
Notes:
id %in% sample.id
***

### Looking at Samples of Households

```{r Looking at Sample of Households}

set.seed(4230)
sample.id <- sample(yo$id, 16)
ggplot(data = subset(yo, id %in% sample.id), aes(x = time, y = price)) + 
  facet_wrap( ~ id) + 
  geom_line() + 
  geom_point(aes(size = all.purchase), pch = 1)
  
### how many yogurt purchased for each househould
library(ggplot2)
library(dplyr)
yo.count <- yo %>% 
  group_by(id) %>% 
  summarise(yo.purchase = sum(all.purchase))

ggplot(data = yo, aes( x = id, y = all.purchase)) + 
  geom_bar(stat = 'summary', fun.y = sum)


```

***

### The Limits of Cross Sectional Data
Notes:

***
```{r}

install.packages('GGally')
library(GGally)

set.seed(1836)
pf_subset <- pf[, c(2:15)]
ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000),])

```

### Many Variables
Notes:

***
```{r}

nci <- read.table('nci.tsv')
colnames(nci) <- c(1:64)
library(reshape2)
nci.long.samp <- melt(as.matrix(nci[1:200, ]))
names(nci.long.samp) <- c('gene', 'case', 'value')
head(nci.long.samp)
```

### Scatterplot Matrix
Notes:

***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

Lesson 5

Here we take many basic techniques you learned in previous lesson and extended them to look at partterns across many variables at once. We started with simple extensions to the scatterplot, and plots of conditional summaries that you worked with in lesson 4, such as adding multiple summaries for multiple groups.Then, we tried some techniques for examining a large number of variables at once, such as scatter-plot matrices and heatmaps. We also learned how to reshape data from broad data with one row per case to aggregate data with one row per combination of variables, and we moved back and forth between long and wide data formats.

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

